MS Prelim Data K23

### 2024-05-28 scripts/first_pass_adapting_K23_analysis_20231031.Rmd
adding yeo things

```{r read_in_fascicle_info}
##########################################################
#             Read in Fascicle info             #
##########################################################

@@

#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])

fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77, drop cranial nerves

########
#fascicle volumes in depression network
########
fascicle_volumes_dep <- read.csv("/Users/eballer/BBL/msdepression/results/streamline_volume_within_dep_network.csv", header = T, sep = ",")

#add column to indicate whether does not meet criteria for inclusion in dep mask (0), any overlapping voxels (1), or >10% overlapping voxels (2)
fascicle_volumes_dep_network <- fascicle_volumes_dep %>%
  replace(is.na(.), 0) %>%
  mutate(inDepMask=case_when(prop_in_mask == 0 ~ 0, (prop_in_mask > 0 & prop_in_mask < 0.1) ~ 1, prop_in_mask >= 0.1 ~ 2))

#write.table(fascicle_volumes_dep_network, paste0(homedir,  "/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping.csv"), sep = ",", col.names=T)#, row.names = F, quote = F, col.names = T)

fascicle_names_nondep_network<- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 0)]

fascicle_names_dep_network_all <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask > 0)]

#fascicle_mapping, returns numerical mapping as well as names of tracts, and whether in or out of dep network
fascicle_bundle_mapping <- get_fascicle_bundle_mapping()

##########################################################
#             Merge with data_empi              #
##########################################################

@@

#lm
fascicle_number_of_anxiety_dx_lm <- lapply(fascicle_names_no_cranial_nerves, function(x) 
{
  lm(substitute(i ~ number_of_anxiety_comorbidities_extended, list(i = as.name(x))), data = df_demo_and_fascicles)
})
names(fascicle_number_of_anxiety_dx_lm) <- fascicle_names_no_cranial_nerves
fascicle_number_of_anxiety_dx_anova <- lapply(fascicle_number_of_anxiety_dx_lm, anova)

#lm
fascicle_depANDanx_dx_lm <- lapply(fascicle_names_no_cranial_nerves, function(x) 
{
  lm(substitute(i ~ Has.depANDanx.dx, list(i = as.name(x))), data = df_demo_and_fascicles)
})
names(fascicle_depANDanx_dx_lm) <- fascicle_names_no_cranial_nerves
fascicle_depANDanx_dx_anova <- lapply(fascicle_depANDanx_dx_lm, anova)

#lm
fascicle_number_of_psychiatric_dx_lm <- lapply(fascicle_names_no_cranial_nerves, function(x) 
{
  lm(substitute(i ~ number_of_psychiatric_comorbidities_extended, list(i = as.name(x))), data = df_demo_and_fascicles)
})
names(fascicle_number_of_psychiatric_dx_lm) <- fascicle_names_no_cranial_nerves

#fdr corrected
fascicle_number_of_psychiatric_dx_anova_fdr <- fdr_anova_generic(fascicle_number_of_psychiatric_dx_anova, 1)
print(fascicle_number_of_psychiatric_dx_anova_fdr)
```

# Yeo network analysis attempts
###Read in fascicle info
```{r read_in_fascicle_info}
##########################################################
#             Read in Fascicle info             #
##########################################################
########
#fascicle volumes in yeo networks
########

@@

#read in file with binarized yes/no 
fascicle_volumes_yeo_7_binarized <- read.csv(paste0(homedir_msdep, "results/yeo_7_network_overlap_proportions_binarized.txt"), header = T, sep = ",")

@@

#check that order of df is same as yeo networks
new_df <- gsub(pattern = "_vol", replacement = "", x = names(df_demo_and_fascicles_prop_and_vol[grep(pattern = "_vol", x = names(df_demo_and_fascicles_prop_and_vol))]))

@@

df_w_yeo_volume_info <- dep_and_healthy_groups_for_ICD_analysis_with_hydra_prop %>%
  filter(Hydra_k2 != -1)  %>% 
 # mutate(total_fascicle_vol_lost = rowMeans(across(all_of(fascicle_names_w_vol_suffix)))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  dplyr::select(EMPI, Hydra_k2, ends_with("_vol")) %>%
  ungroup()

@@

#this has been checked with eyeballs and works. You CAN trust the df_w_yeo_vol_info
for (network in yeo_network_names) {

  for (subj in 1:dim(df_w_yeo_volume_info)[1]) {
      # multiply the binary vector of the network with the fascicle volumes of injury, sum, and store in data frame
      text_to_eval <- paste0("df_w_yeo_volume_info$", network,"[", subj,"]=sum(fascicle_volumes_yeo_7_binarized_no_cranial_nerves$", network,
      "*df_w_yeo_volume_info[",subj,",2:(dim(df_w_yeo_volume_info)[2] - 7)])")
      eval(parse(text=text_to_eval))
  }
}

@@

>>>>

### Read in fascicle info, use this to determine proportion and volume of injury within each fascicle, and save it in the df_demo_and_fascicles_df
```{r read_in_fascicle_info}

### by the end of this chunk, the df should include 77 columns each for proportion of injury and volume of injury in each fascicle

##########################################################
#             Read in Fascicle info             #
##########################################################

@@

## Merge data with good mimosa
```{r merge_with_good_mimosa}
##########################################################
#             Merge with data_empi              #
##########################################################

@@

```{r add_suffix_to_each_measure}

#########################################
## Get vectors with names of fascicles ##
#########################################

#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77, drop cranial nerves

fascicle_names_w_prop_suffix <- gsub(pattern = "(.*)", replacement = "\\1_prop", perl=TRUE, x = fascicle_names_no_cranial_nerves)
fascicle_names_w_vol_suffix <- gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE, x = fascicle_names_no_cranial_nerves)

#read in volumes of fascicles and make columns reflecting volume of disease injured per fascicle

##########################################################
#  Add columns for volume and change proportion names    #
##########################################################

#read in fiber volume values
fiber_volume_values <- read.csv("/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv", sep=",", header = T)

#remove cranial nerves
fiber_volume_values_no_cranial_nerves <- fiber_volume_values %>% filter(fascicle %in% fascicle_names_no_cranial_nerves) %>%
  slice(order(factor(fascicle, levels = fascicle_names_no_cranial_nerves)))

#extract data frame that only includes the fascicle proportion lost
df_fascicles_prop <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names_no_cranial_nerves)] 
df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names_no_cranial_nerves)]

#multiply each row by the vector of total volumes to extract volume of disease within fascicle
df_fascicles_volumes <- sweep(x = df_fascicles_prop, MARGIN = 2,                        # Sweep with Complex STATS
                  STATS = fiber_volume_values_no_cranial_nerves$volume_full, FUN = "*") %>%
  mutate_if(is.numeric, round) #round all values

# change name of df_fascicle_volumes to have the suffix, and add a suffix for the proportion measures
names(df_fascicles_prop) <- fascicle_names_w_prop_suffix
names(df_fascicles_volumes) <- fascicle_names_w_vol_suffix

#combine back with data_fram
#df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
df_demo_and_fascicles <- cbind(df_everything_else, df_fascicles_prop, df_fascicles_volumes)
```

# Yeo network analysis attempts
###Read in fascicle info
```{r read_in_yeo_fascicle_info}
##########################################################
#             Read in Fascicle info             #
##########################################################
########
#fascicle volumes in yeo networks
########

@@

#read in file with binarized yes/no 
fascicle_volumes_yeo_7_binarized <- read.csv("/Users/eballer/BBL/msdepression/results/yeo_7_network_overlap_proportions_binarized.txt", header = T, sep = ",")

@@

#check that order of df is same as yeo networks
new_df <- gsub(pattern = "_vol", replacement = "", x = names(df_demo_and_fascicles[grep(pattern = "_vol", x = names(df_demo_and_fascicles))]))

@@

df_w_yeo_volume_info <- df_demo_and_fascicles %>%
  # mutate(total_fascicle_vol_lost = rowMeans(across(all_of(fascicle_names_w_vol_suffix)))) %>%
  dplyr::select(EMPI, ends_with("_vol")) %>%
  ungroup()

@@

#this has been checked with eyeballs and works. You CAN trust the df_w_yeo_volume_info
for (network in yeo_network_names) {

  for (subj in 1:dim(df_w_yeo_volume_info)[1]) {
      # multiply the binary vector of the network with the fascicle volumes of injury, sum, and store in data frame
      text_to_eval <- paste0("df_w_yeo_volume_info$", network,"[", subj,"]=sum(fascicle_volumes_yeo_7_binarized_no_cranial_nerves$", network,
      "*df_w_yeo_volume_info[",subj,",3:(dim(df_w_yeo_volume_info)[2] - 7)])")
     # print(text_to_eval)
      eval(parse(text=text_to_eval))
  }
}

@@

### rename the whole new data frame df_demo_and_fascicles
df_demo_and_fascicles <- df_w_yeo_volume_info_with_prop_of_network %>% 
  dplyr::select(-contains('_vol')) %>% #removes all columns with volume, leaving only the exam date, empi, and yeo values
  right_join(df_demo_and_fascicles, by = c("EXAM_DATE", "EMPI")) #combine with demo_and_fascicles

#summarize means across all fascicles
```{r make_summary_measures_of_fascicle_burden}

df_demo_and_fascicles <- df_demo_and_fascicles %>%
  mutate(total_fascicle_prop_lost_no_cranial_nerves = rowMeans(across(fascicle_names_w_prop_suffix))) %>%
  mutate(total_fascicle_volume_lost_no_cranial_nerves = rowMeans(across(fascicle_names_w_vol_suffix)))
```

#lm
fascicle_number_of_anxiety_dx_lm <- lapply(fascicle_names_w_prop_suffix, function(x) 
{
  lm(substitute(i ~ number_of_anxiety_comorbidities_extended, list(i = as.name(x))), data = df_demo_and_fascicles)
})
names(fascicle_number_of_anxiety_dx_lm) <- fascicle_names_w_prop_suffix
fascicle_number_of_anxiety_dx_anova <- lapply(fascicle_number_of_anxiety_dx_lm, anova)

## limbic area disease by anxiety diagnosis

yeo_lim_no_anxiety <- lm(LIM~number_of_anxiety_comorbidities_extended, data = df_demo_and_fascicles) #nothing

#lm
fascicle_depANDanx_dx_lm <- lapply(fascicle_names_w_prop_suffix, function(x) 
{
  lm(substitute(i ~ Has.depANDanx.dx, list(i = as.name(x))), data = df_demo_and_fascicles)
})
names(fascicle_depANDanx_dx_lm) <- fascicle_names_w_prop_suffix
fascicle_depANDanx_dx_anova <- lapply(fascicle_depANDanx_dx_lm, anova)

#lm
fascicle_number_of_psychiatric_dx_lm <- lapply(fascicle_names_w_prop_suffix, function(x) 
{
  lm(substitute(i ~ number_of_psychiatric_comorbidities_extended, list(i = as.name(x))), data = df_demo_and_fascicles)
})
names(fascicle_number_of_psychiatric_dx_lm) <- fascicle_names_w_prop_suffix

#fdr corrected
fascicle_number_of_psychiatric_dx_anova_fdr <- fdr_anova_generic(fascicle_number_of_psychiatric_dx_anova, 1)
print(fascicle_number_of_psychiatric_dx_anova_fdr)

## yeo
#lm
psych_comorbidity_lm_yeo <- lapply(yeo_network_names, function(x) 
{
  lm(substitute(i ~ number_of_psychiatric_comorbidities_extended, list(i = as.name(x))), data = anxiety_vs_no_psych)
})
names(psych_comorbidity_lm_yeo) <- yeo_network_names

psych_comorbidity_yeo_anova <- lapply(psych_comorbidity_lm_yeo, anova)

#fdr corrected
psych_comorbidity_yeo_anova_fdr <- fdr_anova_generic(psych_comorbidity_yeo_anova, 1)
print(psych_comorbidity_yeo_anova_fdr)

##compare disease in anxiety versus patients with no psychiatric comorbidity
```{r anxiety_vs_healthy_ish}

anxiety_vs_no_psych <- df_demo_and_fascicles %>%
  mutate(anxietyGroupVar = ifelse(Has.anxietydx, 2, ifelse(healthy_ish, 1, 0))) %>% #complicated, if they have an anxiety disorder, they get a 2, if they are health-ish, no psych dx or meds, 1, otherwise, exclude
  filter(anxietyGroupVar != 0) %>% #get rid of the people who are not anxious and not healthy_ish, n =650, lost n=173
  group_by(EMPI) %>%
  arrange(EXAM_DATE) %>%
  slice(1) %>%
  ungroup

#lm
anxiety_vs_no_psych_lm <- lapply(fascicle_names_w_prop_suffix, function(x) 
{
  lm(substitute(i ~ anxietyGroupVar, list(i = as.name(x))), data = anxiety_vs_no_psych)
})
names(anxiety_vs_no_psych_lm) <- fascicle_names_w_prop_suffix

anxiety_vs_no_psych_lm_anova <- lapply(anxiety_vs_no_psych_lm, anova)

#fdr corrected
anxiety_vs_no_psych_lm_anova_fdr <- fdr_anova_generic(anxiety_vs_no_psych_lm_anova, 1)
print(anxiety_vs_no_psych_lm_anova_fdr)

## yeo
#lm
anxiety_vs_no_psych_lm_yeo <- lapply(yeo_network_names, function(x) 
{
  lm(substitute(i ~ anxietyGroupVar, list(i = as.name(x))), data = anxiety_vs_no_psych)
})
names(anxiety_vs_no_psych_lm_yeo) <- yeo_network_names

anxiety_vs_no_psych_yeo_anova <- lapply(anxiety_vs_no_psych_lm_yeo, anova)

#fdr corrected
anxiety_vs_no_psych_yeo_anova_fdr <- fdr_anova_generic(anxiety_vs_no_psych_yeo_anova, 1)
print(anxiety_vs_no_psych_yeo_anova_fdr)
